---
title: "POS Example"
author: "Dylan Hyder"
date: "November 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This lesson we will work with 4 POS Taggers:

- openNLP
- tagger by trinker
- RDRPOSTagger by bnosac

# Install & Load Necessary Packages

First lets install/ load packages for openNLP
```{r}
#if (!require("pacman")) install.packages("pacman")

pacman::p_load(NLP,openNLP)

```

Next lets install/ load packages for tagger
```{r}
pacman::p_load_gh(c(
    "trinker/termco", 
    "trinker/coreNLPsetup",        
    "trinker/tagger"
))

pacman::p_load(tidytext)
```

Last, lets install/ load packages for RDRPOSTagger
```{r}
pacman::p_load(devtools)

pacman::p_load_gh("bnosac/RDRPOSTagger")


#devtools::install_github("bnosac/RDRPOSTagger")
#devtools::install_github("ropensci/tokenizers")
#library("RDRPOSTagger")
#library("tokenizers")
#library("koRpus")
```

Now lets load in some text to tag for parts of speech. Make sure they are formatted as strings.
```{r}

text <- "Rubik's Cube is a 3-D combination puzzle invented in 1974 by Hungarian sculptor and professor of architecture Ernő Rubik. Originally called the Magic Cube, the puzzle was licensed by Rubik to be sold by Ideal Toy Corp. in 1980 via businessman Tibor Laczi and Seven Towns founder Tom Kremer, and won the German Game of the Year special award for Best Puzzle that year. As of January 2009, 350 million cubes had been sold worldwide making it the world's top-selling puzzle game. It is widely considered to be the world's best-selling toy."

```

Now lets see how to tag these sentences with each package...

# openNLP

For information about the openNLP package go to: https://cran.r-project.org/web/packages/openNLP/openNLP.pdf

openNLP prefers working with text in string format (at least from what we experienced.) Lets make sure our text is formatted correctly.
```{r}
text <- as.String(text)  # Ensure text is in string format 
```

First, we need to break up the text into sentences and words. Do this by calling a sentence and word token annotator through openNLP. Notice the methodologies being used in the sentence and word token annotator.
```{r}
sent_token_annotator <- Maxent_Sent_Token_Annotator()
word_token_annotator <- Maxent_Word_Token_Annotator()

sent_token_annotator
word_token_annotator
```

Now lets format the text to identify all words and sentences. Notice how openNLP organizes the text input.
```{r}
WordAndSent_AnnText <- annotate(text,list(sent_token_annotator,word_token_annotator))
WordAndSent_AnnText
```

Now that the text is formatted into words and sentences. Lets tag each word with its POS. First we must call a POS annotator. Notice the methodology that is being used.
```{r}
pos_tag_annotator <- Maxent_POS_Tag_Annotator()
pos_tag_annotator
```

Another option for a POS annotator in **openNLP** uses a Perceptron Neural Net method. Check out the following link to find out more info. http://opennlp.sourceforge.net/models-1.5/

Lets annotate the text with the Maxent method
```{r}
WordAndSent_AnnText <- annotate(text,pos_tag_annotator, WordAndSent_AnnText)
WordAndSent_AnnText
```

Notice how each word has been labeled to a part of speech. However, it is not very useful in this format because we can't see what each word is. Let's fix that... First lets save only the words and drop the sentences.
```{r}
WordAndSent_AnnText <- subset(WordAndSent_AnnText,type=="word") 
WordAndSent_AnnText
```

Next lets identify all the tags that were used...
```{r}
tags <- sapply(WordAndSent_AnnText$features,'[[',"POS")
tags
```

Perhaps we want to know the number of times each POS tag occurred.
```{r}
table(tags)
```

Finally, lets augment the text with the POS tags so we can see it all together. Notice the format of the output this produces.
```{r}
sprintf("%s/%s",text[WordAndSent_AnnText],tags)

```

# trinker\tagger

## Using openNLP

For information about this package go to: https://github.com/trinker/tagger

This package is essentially a wrapper for openNLP and NLP which makes POS tagging *much* easier!! 

openNLP prefers working with text in character format (at least from what we experienced.) Lets make sure our text is formatted correctly.
```{r}
text <- as.character(text)  # Ensure text is in string format 
```

Now lets tag the text with POS...
```{r}
text_w_POS <- tag_pos(text, engine = "openNLP")
text_w_POS
```

Much easier, isn't it. To figure out what is going under the hood type *?tag_pos*. Also, do you notice how the final output is in a different form than openNLP?

## Using coreNLP

First make sure that you have **coreNLP** installed. T. Rinker makes this easy with the function *check_setup*: 
```{r}
#check_setup()
```

If coreNLP is installed lets tag the text using the coreNLP engine instead of the openNLP engine
```{r}
#tag_pos(text, engine = "coreNLP")
```


# RDRPOSTagger

First lets define the POS tagger model
```{r}
unipostagger <- rdr_model(language = "English", annotation = "POS")           #define model for RDRPOSTagger
unipostagger2 <- rdr_model(language = "English", annotation = "UniversalPOS") #define model for RDRPOSTagger
```

Now that the POS tagger is defined lets tag our text
```{r}
rdr_pos(unipostagger, text)
```

```{r}
rdr_pos(unipostagger2, text)
```


# Visualizing POS

In this section lets answer two questions:

(1) Is tagger with the openNLP engine the same as POS tagging with openNLP
(2) How can we easily visualize POS tags

Plots are always nice...

Lets plot the POS tags from openNLP
```{r}
table(tags) 
```


Lets plot the POS tags from tagger
```{r}
#text_w_POS %>%
#  plot()
```

```{r}

```

# **CALM** Example from Yesterday

Last class we had three sentences with the word **calm** that had a different POS tag for each instance. How well does each tagger do at recognizing the difference?

First lets pull in the sentences...
```{r}
csent1 <- "The calm lasted for three days."
csent2 <- "Calm words show quiet minds."
csent3 <- "Calm your angry friend."

sent1 <- as.String(csent1)
sent2 <- as.String(csent2)
sent3 <- as.String(csent3)

```

Next lets tag each sentence using openNLP, tagger, and RDRPOSTagger.
```{r}
openNLP_sent1 <- annotate(sent1,list(sent_token_annotator,word_token_annotator))
openNLP_sent1 <- annotate(sent1,pos_tag_annotator, openNLP_sent1)
openNLP_sent1 <- subset(openNLP_sent1,type=="word") 
openNLP_sent1_tags <- sapply(openNLP_sent1$features,'[[',"POS")

openNLP_sent2 <- annotate(sent2,list(sent_token_annotator,word_token_annotator))
openNLP_sent2 <- annotate(sent2,pos_tag_annotator, openNLP_sent2)
openNLP_sent2 <- subset(openNLP_sent2,type=="word") 
openNLP_sent2_tags <- sapply(openNLP_sent2$features,'[[',"POS")

openNLP_sent3 <- annotate(sent3,list(sent_token_annotator,word_token_annotator))
openNLP_sent3 <- annotate(sent3,pos_tag_annotator, openNLP_sent3)
openNLP_sent3 <- subset(openNLP_sent3,type=="word") 
openNLP_sent3_tags <- sapply(openNLP_sent3$features,'[[',"POS")


RDRtags_sent1 <- rdr_pos(unipostagger, csent1)
RDRtags_sent2 <- rdr_pos(unipostagger, csent2)
RDRtags_sent3 <- rdr_pos(unipostagger, csent3)

tagger_openNLP_sent1 <- tag_pos(csent1, engine = "openNLP")
tagger_openNLP_sent2 <- tag_pos(csent2, engine = "openNLP")
tagger_openNLP_sent3 <- tag_pos(csent3, engine = "openNLP")

#tagger_coreNLP_sent1 <- tag_pos(sent1, engine = "coreNLP")
#tagger_coreNLP_sent2 <- tag_pos(sent2, engine = "coreNLP")
#tagger_coreNLP_sent3 <- tag_pos(sent3, engine = "coreNLP")

```

Lets compare how they do...



- Sentence 1:
```{r}
sprintf("%s/%s",sent1[openNLP_sent1],openNLP_sent1_tags)
tagger_openNLP_sent1
RDRtags_sent1
#tagger_coreNLP_sent1
```

- Sentence 2:
```{r}
sprintf("%s/%s",sent2[openNLP_sent2],openNLP_sent2_tags)

tagger_openNLP_sent2
RDRtags_sent2
#tagger_coreNLP_sent2
```

- Sentence 3:
```{r}
sprintf("%s/%s",sent3[openNLP_sent3],openNLP_sent3_tags)

tagger_openNLP_sent3
RDRtags_sent3
#tagger_coreNLP_sent3
```

